### 2022-04-12 Analysis/sup_mat.Rmd
Tagu initial selections

---
title: 'SVG: The Paper'
author: "Alasdair and Anna"
date: "11/04/2022"
output: html_document
---

```{r}
library(tidyverse)
library(patchwork)
```

# Initial Selection

## QJEP Dataset

```{r}
d <- read_csv("data/clarke_2020_qjep.csv", show_col_types = FALSE) %>%
  filter(found == 1) %>%
  mutate(condition = as_factor(condition))

ggplot(d, aes(x, y, colour = condition)) + geom_point(alpha = 0.25) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  facet_wrap(~observer)

d %>% group_by(observer, condition) %>% 
  summarise(x = median(x), y = median(y)) %>%
  ggplot(aes(x, y, colour = condition)) + geom_point(alpha = 0.25) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))

plt_x <- ggplot(d, aes(x)) + geom_histogram(bins = 10) 
plt_y <- ggplot(d, aes(y)) + geom_histogram(bins =  8) 
plt_x + plt_y
```

## Tagu dataset

```{r}
d <- read_csv("data/tagu_2020_prox_mouse.csv", show_col_types = FALSE) %>%
  filter(found == 1) %>%
  mutate(condition = as_factor(condition))

ggplot(d, aes(x, y, colour = condition)) + geom_point(alpha = 0.25) +
  #coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  facet_wrap(~observer)

d %>% group_by(observer, condition) %>% 
  summarise(x = median(x), y = median(y)) %>%
  ggplot(aes(x, y, colour = condition)) + geom_point(alpha = 0.25) #+
  #coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))

plt_x <- ggplot(d, aes(x)) + geom_histogram(bins = 10) 
plt_y <- ggplot(d, aes(y)) + geom_histogram(bins =  8) 
plt_x + plt_y
```