### 2022-04-12 Analysis/sup_mat.Rmd
setting up Rmd etc

---
title: 'SVG: The Paper'
author: "Alasdair and Anna"
date: "11/04/2022"
output: html_document
---

```{r}
library(tidyverse)
library(patchwork)
library(tidybayes)
```

```{r}
# set ggplot2 options
theme_set(theme_minimal())

our_cols <- c("#1b9e77", "#7570b3")
options(ggplot2.discrete.fill = our_cols ,
        ggplot2.discrete.colour = our_cols)
```

# Trial Level Posterior Predictions

```{r echo = FALSE}
source("functions/get_model_params.R")
source("functions/compute_weights.R")
```

## Clarke et al (2022)

```{r}
a <- readRDS("scratch/qjep_model_weights.rda")

a %>% group_by(condition, observer, found) %>%
  summarise(meanb = mean(b),
            prop_best = mean(selected_max), .groups = "drop") %>%
  filter(found > 1) %>%
  mutate(chance = 1/(41-found))-> a_agg
```

```{r, fig.cap = "(Left) The average weight assigned to each selected target by our model. (Right) The proportion of trials the item with the largest assigned weight as selected by the participant.  Each dot shows data from an individual particpant in a condition and the shaded region indicates the invertal in which we expect 67% of participants to fall."}

# plot target selected weights
ggplot(a_agg, aes(x = found, y = meanb, colour = condition, fill = condition)) + 
  geom_jitter(data = filter(a_agg, found<40), width = 0.1, height = 0, alpha = 0.2) + 
  stat_lineribbon(.width = 0.67, alpha = 0.50) +
  geom_path(data = filter(a_agg, observer == 1, condition == "feature"), 
                          aes(y = chance), linetype = 2, colour = "black") + 
  geom_point(data = tibble(x=40, y=1), aes(x, y), size = 1.5, colour = "black", fill = "grey") + 
  scale_x_continuous(breaks = c(2, 10, 20, 40), "target selection", expand = c(0.01, 0.01)) + 
  scale_y_continuous("average weight from model", expand = c(0.01, 0.01)) -> plt_b

ggplot(a_agg, aes(x = found, y = prop_best, colour = condition, fill = condition)) + 
  geom_jitter(data = filter(a_agg, found<40), width = 0.1, height = 0.00, alpha = 0.2) + 
  stat_lineribbon(.width = 0.67, alpha = 0.50) +
  geom_path(data = filter(a_agg, observer == 1, condition == "feature"), 
            aes(y = chance), linetype = 2, colour = "black") + 
  geom_point(data = tibble(x=40, y=1), aes(x, y), size = 1.5, colour = "black", fill = "grey") + 
  scale_x_continuous(breaks = c(2, 10, 20, 40), "target selection", expand = c(0.01, 0.01)) + 
  scale_y_continuous("proportion most likely was selected", expand = c(0.01, 0.01)) -> plt_c

plt_b + plt_c + plot_layout(guides = "collect")  &
  theme(legend.position = 'bottom',
        legend.direction = 'horizontal')
ggsave("../Figures/qjep_preds.png", width = 9, height = 4)
```

# Initial Selection

## QJEP Dataset

```{r}
d <- read_csv("data/clarke_2020_qjep.csv", show_col_types = FALSE) %>%
  filter(found == 1) %>%
  mutate(condition = as_factor(condition))

ggplot(d, aes(x, y, colour = condition)) + geom_point(alpha = 0.25) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  facet_wrap(~observer)

d %>% group_by(observer, condition) %>% 
  summarise(x = median(x), y = median(y)) %>%
  ggplot(aes(x, y, colour = condition)) + geom_point(alpha = 0.25) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))

plt_x <- ggplot(d, aes(x)) + geom_histogram(bins = 10) 
plt_y <- ggplot(d, aes(y)) + geom_histogram(bins =  8) 
plt_x + plt_y
```

## Tagu dataset

```{r}
d <- read_csv("data/tagu_2020_prox_mouse.csv", show_col_types = FALSE) %>%
  filter(found == 1) %>%
  mutate(condition = as_factor(condition))

ggplot(d, aes(x, y, colour = condition)) + geom_point(alpha = 0.25) +
  #coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) + 
  facet_wrap(~observer)

d %>% group_by(observer, condition) %>% 
  summarise(x = median(x), y = median(y)) %>%
  ggplot(aes(x, y, colour = condition)) + geom_point(alpha = 0.25) #+
  #coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))

plt_x <- ggplot(d, aes(x)) + geom_histogram(bins = 10) 
plt_y <- ggplot(d, aes(y)) + geom_histogram(bins =  8) 
plt_x + plt_y
```