### 2022-05-04 Analysis/index_prpic.Rmd
Starting work on Ppric

---
title: 'SVG: The Paper'
author: "Alasdair Clarke, Amelia Hunt &  Anna Hughes"
date: "13/04/2022"
output:
  tufte::tufte_html:
    number_sections: true
    fig_height: 4
    bibliography: literature.bib
    link-citations: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits = 3)
set.seed(2022)
```

```{r, message=FALSE}
# attach packages
library(tidyverse)
library(patchwork)
library(tidybayes)
library(ggrepel)
library(truncnorm)
```

```{r}
# set ggplot2 options
theme_set(ggthemes::theme_tufte() + theme(plot.background = element_rect(fill = "#fffff8", colour = "#fffff8")))

# some more colours
col3 = c("#d04e00", "#ddc000","#34b6c6",  "#79ad41", "#4063a3")

our_cols <- c("#1b9e77", "#7570b3")
options(ggplot2.discrete.fill = our_cols ,
        ggplot2.discrete.colour = our_cols)
```

Terminology questions:
- item v target
- participant v observer

# Trial Level Posterior Predictions{#label}

```{r echo = FALSE}
source("functions/get_model_params.R")
source("functions/compute_weights.R")
```

We will use the visual foraging data from @clarke2022 to assess the model's trial level performance. As the current version of our model only takes the $p_A$ bias into account for the initial target selection, we will ignore these data points for now.

```{r, echo = TRUE}
# read in data
d <- read_csv("data/squirrel_spatial.csv", show_col_types = FALSE) %>%
  mutate(condition = as_factor(condition),
         condition = fct_recode(condition, feature = "1", conjunction = "2"),
         targ_type = as_factor(targ_type))
```

We will use the model fit as detailed in @clarke2020compbio. For simplicity sake, we will characterise all of our estimated posterior probability densities with just the mean value. This gives us four parameters per participant: $b_S, b_A, b_P, b_D$.

Using these fitted values, we can now step through each target selection in the dataset and computer the probability weights our model assigns to each remaining target. For each target selection, we record:

- the item assigned the highest weight by our model.
- whether this item was then selected by the human participant.

```{r echo = TRUE}
# item weights pre-computed - see scripts/extract_item_weights_using_model.R
a <- readRDS("scratch/prpic_model_weights.rda") %>%
  filter(found != 1) # remove initial selections

# compute the average weight for each participant x condition x item selection
# also compute the proportion of times in which the item selected by the 
# participant was judged the most likely to be selected by our model.
a %>% group_by(condition, observer, found) %>%
  summarise(meanb = mean(b),
            prop_best = mean(selected_max), .groups = "drop") %>%
  # finally, also compute what we would expect under a null-model
  mutate(chance = 1/(41-found))-> a_agg
```

We can then plot this data to show how our ability to predict which target will be selected next varies throughout a trial. We can see a difference between *feature* and *conjunction* foraging, with *feature* foraging being more predictable.

```{r, fig.cap = "(*left*) The average weight assigned to each selected target by our model. (*right*) The proportion of trials the item with the largest assigned weight was selected by the participant.  Each dot shows data from an individual participant in a condition and the shaded region indicates the interval in which we expect 67% of participants to fall."}

# plot target selected weights
ggplot(a_agg, aes(x = found, y = meanb, colour = condition, fill = condition)) + 
  geom_jitter(data = filter(a_agg, found<40), width = 0.1, height = 0, alpha = 0.2) + 
  stat_lineribbon(.width = 0.67, alpha = 0.50) +
  geom_path(data = filter(a_agg, observer == 1, condition == "feature"), 
                          aes(y = chance), linetype = 2, colour = "black") + 
  geom_point(data = tibble(x=40, y=1), aes(x, y), size = 1.5, colour = "black", fill = "grey") + 
  scale_x_continuous(breaks = c(2, 10, 20, 40), "target selection", expand = c(0.01, 0.01)) + 
  scale_y_continuous("average weight from model", expand = c(0.01, 0.01)) -> plt_b

ggplot(a_agg, aes(x = found, y = prop_best, colour = condition, fill = condition)) + 
  geom_jitter(data = filter(a_agg, found<40), width = 0.1, height = 0.00, alpha = 0.2) + 
  stat_lineribbon(.width = 0.67, alpha = 0.50) +
  geom_path(data = filter(a_agg, observer == 1, condition == "feature"), 
            aes(y = chance), linetype = 2, colour = "black") + 
  geom_point(data = tibble(x=40, y=1), aes(x, y), size = 1.5, colour = "black", fill = "grey") + 
  scale_x_continuous(breaks = c(2, 10, 20, 40), "target selection", expand = c(0.01, 0.01)) + 
  scale_y_continuous("proportion most likely was selected", expand = c(0.01, 0.01)) -> plt_c

plt_b + plt_c + plot_layout(guides = "collect")  &
  theme(legend.position = 'bottom',
        legend.direction = 'horizontal')

rm(plt_c)

```

## Calibration

Is our model calibrated? By this we mean, if our model assigns an item a probability $p$ of being selected next, then is it actually selected (by the human participant) this often?

We will calculate this as follows: for each target selection (on each trial, for each participant) we look at the weight assigned to the most likely item, and then look at whether it was selected or not.

```{r, echo = TRUE}
my_breaks <- seq(0.0, 1, 0.05)

a %>% mutate(b_bin = cut(max_b, breaks = my_breaks, labels = FALSE)) %>%
  group_by(condition, b_bin) %>% 
  summarise(acc = mean(selected_max), .groups = "drop") %>%
  mutate(b_bin = as.numeric(b_bin)/(length(my_breaks)-1)) -> a_cut
```

We can see from the Figure below that our model is well-calibrated (at least in terms of the target with the highest assigned weight). If we repeat this calculation for the 2nd and 3rd ranked candidate items we can see that we slightly underweight these items, but we are still pretty accurate.

```{r, fig.height=3.5, fig.cap = "(*left*) Calibration plot for our foraging model. The *x*-axis gives the largest weight assigned by the model while the *y*-axis shows how often that target was actually selected by a human participant. (*right*) This plot shows how often the 2nd and 3rd ranked items are selected based on the weights assigned by the model."}

ggplot(a_cut, aes(b_bin, acc, colour = condition, fill = condition, shape = condition)) + 
  geom_point() + 
  geom_line(stat = "smooth", method = "loess", 
            formula = y ~ x, alpha = 0.65, size = 2, se = F) +
  geom_abline(linetype = 2) +
  scale_x_continuous("largest item weight") +
  scale_y_continuous("proportion of times selected") +
  theme(legend.position = "bottom") -> plt1

a %>% select(max_b2, max_b3, selected_max2, selected_max3) %>%
  unite(`2`, max_b2, selected_max2) %>%
  unite(`3`, max_b3, selected_max3) %>%
  pivot_longer(c(`2`, `3`), names_to = "rank", values_to = "b") %>%
  separate(b, into = c("b", "selected"), sep = "_", convert = TRUE) %>%
  mutate(b_bin = cut(b, breaks = my_breaks, labels = FALSE)) %>%
  group_by(b_bin) %>% 
  summarise(acc = mean(selected), .groups = "drop") %>%
   mutate(b_bin = as.numeric(b_bin)/(length(my_breaks)-1)) %>%
  filter(is.finite(acc)) %>%
  ggplot(aes(b_bin, acc)) + 
  geom_point() + 
  geom_line(stat = "smooth", method = "loess", 
            formula = y ~ x, alpha = 0.65, size = 2, se = F) +
  geom_abline(linetype = 2) +
  scale_x_continuous("runner up item weights") +
  scale_y_continuous("proportion of times selected") -> plt2

 plt1 + plt2

ggsave("../Figures/qjep_preds.pdf", 
       plt_b + plt1 + plt2 + 
         plot_layout(guides = "collect") & theme(legend.position = "bottom"),
       width = 10, height = 4)

rm(plt1, plt2, plt_a)
```

```{r}
rm(a_cut, my_breaks, plt1, plt2)
```

### Individual Differences

How often does each participant select the target with the largest weight?

```{r, fig.height=3.5, fig.cap = "Prediction scores for participants. Boxplots show quartile range and the grey lines indicate individual participants. (The dots indicate outliers.)", echo= TRUE}

a_agg %>% group_by(observer, condition) %>%
  summarise(accuracy = mean(prop_best), .groups = "drop") -> a_acc

ggplot(a_acc, aes(x= condition, y = accuracy, fill = condition)) + 
  geom_boxplot() +
  geom_line(aes(group = observer), alpha = 0.25, colour = "sienna4") +
  scale_y_continuous("model accuracy",  
                     limits = c(0.35, 0.8), breaks = seq(0.3, 0.8, 0.1)) -> plt_a

plt_a
```

Is this explained by differences in proximity weighting? i.e., are participants with weaker proximity biases harder to predict? Yes!

```{r fig.height = 6, fig.cap = "Accuracy of our model varies with the strength of an individual's model parameters. We can see two clear outlier participants (marked with an X). "}
# computed in xxxx.R
fit <- readRDS("scratch/qjep_model_fit.rda") %>%
  full_join(a_acc, by = c("observer", "condition")) %>%
  mutate(is_outlier = if_else(observer %in% c(35, 45), TRUE, FALSE))

fit_plt <- pivot_longer(fit, c(pA, pS, bP, bM), 
                     names_to = "parameter", values_to = "value")

ggplot(fit_plt, aes(value, accuracy, colour = condition)) + 
  geom_point(aes(shape = is_outlier)) +
  geom_smooth(data = filter(fit_plt, is_outlier == 0), method = "lm", formula = y ~ x, se = FALSE) + 
  scale_shape_manual(values = c(19, 4), guide = "none") + 
  facet_wrap(~parameter, scales = "free") +
  theme(legend.position = "bottom")

plt_b <- ggplot(filter(fit_plt, parameter == "bP"), 
                aes(value, accuracy, colour = condition)) + 
  geom_point(aes(shape = is_outlier)) +
  geom_smooth(data = filter(fit_plt, is_outlier == 0, parameter == "bP"), method = "lm", formula = y ~ x, se = FALSE) + 
  scale_shape_manual(values = c(19, 4), guide = "none") + 
  theme(legend.position = "none") + 
   scale_y_continuous("model accuracy",  
                     limits = c(0.35, 0.8), breaks = seq(0.3, 0.8, 0.1))

ggsave("../Figures/qjep_indiv_diff.pdf", 
       plt_a + plt_b + 
         plot_layout(guides = "collect") & theme(legend.position = "right"), width = 8, height = 4)
```

As the two outlier participants appear to be doing something different, not captured especially well by our model, we will remove them from the rest of this analysis.

After removing them, we can see that individual differences in proximity tuning account for nearly all of the differences in predictability from person to person, and also between the *feature* and *conjunction* conditions.

```{r, echo = TRUE}
dm <- filter(fit, is_outlier == 0)

summary(lm(accuracy ~ condition * bP, 
           data = dm))
```

We can see that the difference in the proximity parameter between conditions appears to account for most of the variables in our model's accuracy from one participant to the next.

We can also see if the other parameters have an effect.

```{r echo=TRUE}
summary(lm(accuracy ~ pA + pS + bM + bP, 
           data = dm))
```

Both $p_A$ and $b_M$ also have an effect on our model's accuracy, but note there is only a small increases in $R^2$ from just using $b_P$.

```{r}
rm(dm)
```

## Examples of Trials

Let's score every trial by how well the model captures behaviour, and then plot some different examples.

```{r, echo = TRUE}
a %>% filter(!(observer %in% c(35, 45))) %>%
  group_by(observer, condition, trial) %>% 
  summarise(prop_max = mean(selected_max), .groups = "drop") %>% 
  arrange(desc(prop_max)) %>%
  ungroup() -> a_trl
```

```{r fig.height=3.5, fig.cap = "Histogram of how predictiable individual trials are, defined as how often the model assigned the largest weight to the target that was then selected."}
ggplot(a_trl, aes(prop_max)) + 
  geom_histogram(fill = "#d04e00", breaks = seq(0, 1, 0.025), alpha = 0.5) +
  scale_x_continuous("trial predictability")
```